Contextual Combinatorial Cascading Bandits

### 2016-02-01 contextual combinatorial cascading bandits.tex
update experiments and conclusions

\subsubsection{Disjunctive Objective}

@@

In the above, we formulate the general setting of contextual combinatorial cascading bandit with position discount. This setting generalizes several existing results.

If our setting restricts to the setting of combinatorial cascading bandit \cite{kveton2015combinatorial} which coincides with our section \ref{sec:conj} without contextual information and with position discounts $\gamma_k = 1$, our regret bound is of order $O\left(\frac{L}{f^*}\sqrt{nK}\ln(n)\right)$ which has an additional term $\sqrt{\frac{L\ln(n)}{f^*}}$ than their result. The reason is that the best regret bound for stochastic MAB is $O(\frac{\ln(n)}{\Delta})$, which has been achieved by a lot of methods, while the best regret bound for linear bandits is $O\left(\frac{\ln^2(n)}{\Delta}\right)$ by \cite{abbasi2011improved} and we use the technique of linear bandits to derive a confidence ellipsoid of $\theta_{\ast}$.

If our setting restricts to contextual combinatorial bandits \cite{qin2014contextual}, where there are no triggering and no position discounts, so every base arm in action $\bA_t$ will be observed. The random variable $\bO_t$ of observed base arms should be $|\bA_t|$. The probability of full observation of $A$, $p_{t, A}$, should be $1$, so $p^{\ast} = 1$. And the position discounts $\gamma_k$ are all $1$. Then our regret bound is of the same order with theirs. Notice that if a reward function satisfies our assumptions of monotonicity and Lipschitz continuity, then the reward function satisfies their assumptions with the bound $C$ for Lipschitz continuity be $\sqrt{K}$.

@@

\begin{figure*}
	\centering
	\subfigure[Network 1239, $\gamma=1$]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp1239-T5-k100-1.eps}}
	\subfigure[Network 3356, $\gamma = 0.9$]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp3356-T5-k100-9.eps}}
	\subfigure[Network 3257 ]{
		\includegraphics[height = 3 cm, width = 4 cm]{network4-1.eps}}
	\subfigure[Comparison of different $\gamma$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{network-gamma-with-point.eps}}
	\caption{Network, $d=10$}
	\label{fig:network} % 用于交叉引用
\end{figure*}

@@

We evaluate $C^3$-UCB in three experiments. In the following experiments, we let the position discounts $\gamma_k$ have the form of $\gamma^{k-1}$ for some $\gamma$.

@@

In the first experiment, we compare $C^3$-UCB to CombCascade on synthetic problems. This problem is a contextual cascading bandit with $L=100$ items and $K=4$, where at each time $t$ the agent recommends $K$ items to the user. At first, we randomly choose $\theta \in \RR^{d-1}$ with $\|\theta \|_2 = 1$ and let $\theta_* = (\frac{\theta}{2}, \frac{1}{2})$. Then at each time $t$, we randomly assign $x'_{t,a} \in \RR^{d-1}$ with $\|x_{t,a}\|_2 = 1$ to arm $a$ and use $x_{t,a} = (x'_{t,a}, 1)$ to be the context information for arm $a$. This processing will guarantee the inner product $\theta_*^{\top}x_{t,a} = \frac{\theta^{\top}x + 1}{2} \in [0,1]$. Next we generate the weight for arm $a$ at time $t$ by a random sample from Bernoulli distribution with mean $\theta_*^{\top}x_{t,a}$.

We experiment on four settings. Under each setting, the learning agent chooses a set of $K$ items out of $L$ ground items. The first two are of the disjunctive objective where the learning agent observes from the first one of the chosen $K$ items until first one with weight $1$; the last two are of conjunctive objective where the learning agent observes from the first item until the first one with weight $0$. Notice that with $\gamma \leq 1$, the oracle to select a feasible action is to select the set of $K$ items with highest UCBs; $\gamma < 1$ will force the items in the optimal action list in the decreasing order of their UCBs and we also use this order when $\gamma = 1$. The results are shown in Fig.~\ref{fig:synthetic} and our algorithm outperforms {\it CombCascade} algorithm because they do not make use of the context information.

\subsection{MovieLens Data}

In this experiment, we evaluate $C^3$-UCB on a problem of contextual recommendations. We experiment on {\it MovieLens} dataset \cite{lam2013movie} of 2015.

The Learning problem is formulated as follows. There is a big sparse matrix $A \in \{0,1\}^{N_1 \times N_2}$ where $A(i,j) = 1$ denotes user $i$ has watched movie $j$. We split $A$ as $H + F$ by putting entry-$1$ in $A$ to $H$ and $F$ with probability $\sim \mathrm{Ber}(p)$ for some fixed $p$. We can regard $H$ as known information about history 'what users have watched' and regard $F$ as future criterion. We use $H$ to derive feature vectors of both users and movies by SVD decomposition, $H = USM^{\top}$ where $U = (u_1; ...;u_{N_1})$ and $M = (m_1;...;m_{N_2})$. At every time $t$, use $x_{t,i} = u_{i}m_j^{\top}$ as the context information of base arm $i$ and randomly choose a user $I_t$. And use $w_t(m) = F(I_t,m)$ as the weight of movie $m$ at time $t$.

In this problem, any combination of $K$ movies is a feasible action. Notice that it might have problem if we use $(u_{I_t}, m_j)$, the concatenation of $u_{I_t}$ and $m_j$, as the context information. Denote $\theta$ by $(\theta_u, \theta_m)$. So the expected mean of a base arm is of the form $\theta_u^{\top} u_{I_t} + \theta_m^{\top} m_j$ but the first term $\theta_u^{\top} u_{I_t}$ is fixed at time $t$ no matter what $\theta_u$ is. To recommend a list of movies by our algorithm, it is equivalent to recommend the first $K$ movies with highest UCBs listing in decreasing order. Thus, it is equivalent to rank the scores $\theta_m^{\top} m_j$ for each movie $j$. So to make use of the user's context information, some quadratic terms of user's information $u_i$ and movie's information $m_j$ are needed. We use the convolution $u_{i}m_j^{\top}$ as the context information of movie $j$ at time $t$.

First, we randomly choose $L=200$ movies and recommend $K=4$ movies on each round. We experiment with $\gamma=1$ (no position discount) and $\gamma=0.9$, and compare our algorithm with {\it CombCascade}. The results are shown in Fig.~\ref{fig:movielens} and demonstrate the advantage to involve context information in real applications.

@@

In the last experiment, we evaluate $C^3$-UCB on a problem of network routing with the {\it RocketFuel} dataset \cite{spring2004measuring}.

The ground set $E$ is the set of links in the network and the feasible action set $\cS$ is the set of all paths in the network. Before learning, the environment randomly chooses a $d$-dimensional vector $\theta_* \in [0,1]^d$ and at each time $t$, any link $e$ is assigned with a random $d$-dimensional context vector $x_{t,e}$. Notice here both $\theta$ and $x$ have been processed like section \ref{sec:ExpSyn} so that $\theta_*^{\top}x \in [0,1]$. The weight for each edge $e$ is a sample from Bernoulli distribution with mean $\theta_*^{\top}x$.

At each time $t$, a pair of source and terminal nodes are chosen randomly. The learning agent recommends a feasible path $A = (a_1,...,a_{|A|})$ between $v_s$ and $v_t$ to maximize the expected reward in the conjunctive objective. We experiment with different position discounts and different networks. The results are shown in Fig.~\ref{fig:network}.

We also experiment on the influence of different $\gamma$'s to the regret. In recommendations, the position discounts represent users' satisfaction and the learning agent may not know its true value. Suppose the true position discount $\gamma = 0.9$ and the learning agent sets $\gamma$ by her experience. We experiment on different setting $\gamma \in \{0.8, 0.82, \ldots, 0.98, 1.0\}$ with the true criterion $\gamma = 0.9$. The regrets with different $\gamma$'s are illustrated in Figure~\ref{fig:network}(d). The setting $\gamma$ is more close to the true $\gamma$, the regret is tending to be smaller. This experiment shows when there are position discounts in real application, it is better to represent this information to the model.

>>>>

\subsubsection{Disjunctive Objective}
\label{sec:disj}

@@

In the above, we formulate the general setting of contextual combinatorial cascading bandit with position discounts. This setting generalizes several existing results.

If our setting restricts to the setting of combinatorial cascading bandit \cite{kveton2015combinatorial} which coincides with our Section \ref{sec:conj} without contextual information and with position discounts $\gamma_k = 1$, our regret bound is of order $O\left(\frac{L}{f^*}\sqrt{nK}\ln(n)\right)$ which has an additional term $\sqrt{\frac{L\ln(n)}{f^*}}$ than their result. The reason is that the best regret bound for stochastic MAB is $O(\frac{\ln(n)}{\Delta})$, which has been achieved by a lot of methods, while the best regret bound for linear bandits is $O\left(\frac{\ln^2(n)}{\Delta}\right)$ by \cite{abbasi2011improved} and we use the technique of linear bandits to derive a confidence ellipsoid of $\theta_{\ast}$. A similar comparison exists between our Section \ref{sec:disj} and Section 2.3 of \cite{kveton2015combinatorial}.

If our setting restricts to contextual combinatorial bandits \cite{qin2014contextual}, where there are no triggering and no position discounts, so every base arm in action $\bA_t$ will be observed. The random variable $\bO_t$ of observed base arms should be $|\bA_t|$. The probability of full observation of $A$, $p_{t, A}$, should be $1$, so $p^{\ast} = 1$. And the position discounts $\gamma_k$ are all $1$. Then our regret bound of Theorem \ref{thm:main} is of the same order with theirs. Notice that the Lipschitz bound $C$ in their setting is $\sqrt{K}$.

@@

\begin{figure*}
	\centering
	\subfigure[Network 1239, $\gamma=1$]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp1239-T5-k100-1.eps}}
	\subfigure[Network 3356, $\gamma = 0.9$]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp3356-T5-k100-9.eps}}
	\subfigure[reward on network 1755 with latency, $\gamma = 0.9$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp1755-T300000-gamma9-latency.eps}}
	\subfigure[Comparison of different $\gamma$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{network-gamma-with-point.eps}}
	\caption{Network, $d=10$}
	\label{fig:network} % 用于交叉引用
\end{figure*}

@@

We evaluated $C^3$-UCB in three experiments. In the following experiments, we let the position discounts $\gamma_k$ have the form of $\gamma^{k-1}$ for some $\gamma$.

@@

In the first experiment, we compared $C^3$-UCB to CombCascade on synthetic problems. This problem is a contextual cascading bandit with $L=100$ items and $K=4$, where at each time $t$ the agent recommends $K$ items to the user. At first, we randomly chose $\theta \in \RR^{d-1}$ with $\|\theta \|_2 = 1$ and let $\theta_* = (\frac{\theta}{2}, \frac{1}{2})$. Then at each time $t$, we randomly assigned $x'_{t,a} \in \RR^{d-1}$ with $\|x_{t,a}\|_2 = 1$ to arm $a$ and use $x_{t,a} = (x'_{t,a}, 1)$ to be the context information for arm $a$. This processing will guarantee the inner product $\theta_*^{\top}x_{t,a} = \frac{\theta^{\top}x + 1}{2} \in [0,1]$. Next we generated the weight for arm $a$ at time $t$ by a random sample from Bernoulli distribution with mean $\theta_*^{\top}x_{t,a}$.

We experimented on four settings. Under each setting, the learning agent chooses a set of $K$ items out of $L$ ground items. The first two are of disjunctive objective where the learning agent observes a prefix of the chosen $K$ items until the first one with weight $1$; the last two are of conjunctive objective where the learning agent observes from the first item until the first one with weight $0$. Notice that with $\gamma \leq 1$, the oracle selects a set of $K$ items with highest UCBs in their decreasing order. The results are shown in Fig.~\ref{fig:synthetic} and our algorithm outperforms {\it CombCascade} algorithm because they do not make use of the context information.

\subsection{Movie Recommendation}

In this experiment, we evaluated $C^3$-UCB on movie recommendation with data set {\it MovieLens} \cite{lam2013movie} of 2015.

The Learning problem is formulated as follows. There is a big sparse matrix $A \in \{0,1\}^{N_1 \times N_2}$ where $A(i,j) = 1$ denotes user $i$ has watched movie $j$. Next, we split $A$ as $H + F$ by putting $1$-entry in $A$ to $H$ and $F$ with probability $\sim \mathrm{Ber}(p)$ for some fixed $p$. We regard $H$ as known information about history "what users have watched" and regard $F$ as future criterion. We use $H$ to derive feature vectors of both users and movies by SVD decomposition, $H = USM^{\top}$ where $U = (u_1; ...;u_{N_1})$ and $M = (m_1;...;m_{N_2})$. At every time $t$, use $x_{t,i} = u_{i}m_j^{\top}$ as the context information of base arm $i$ and randomly choose a user $I_t$. And use $w_t(m) = F(I_t,m)$ as the weight of movie $m$ at time $t$.

In this problem, any combination of $K$ movies is a feasible action. Notice that it might have problem if we use $(u_{I_t}, m_j)$, the concatenation of $u_{I_t}$ and $m_j$, as the context information. Denote $\theta$ by $(\theta_u, \theta_m)$. So the expected mean of a base arm is of the form $\theta_u^{\top} u_{I_t} + \theta_m^{\top} m_j$ but the first term $\theta_u^{\top} u_{I_t}$ is fixed at time $t$ no matter what $\theta_u$ is. To recommend a list of movies by our algorithm, it is equivalent to recommend the first $K$ movies with highest UCBs listing in decreasing order. Thus, it is equivalent to rank the scores $\theta_m^{\top} m_j$ for each movie $j$. So to make use of the user's context information, some nonlinear terms of user's information $u_i$ and movie's information $m_j$ are needed. To address this, we used the outer product $u_{i}m_j^{\top}$ as the context information of movie $j$ at time $t$.

First, we randomly chose $L=200$ movies and recommended $K=4$ movies on each round. We experimented with both $\gamma=1$ (no position discount) and $\gamma=0.9$, and compared our algorithm with {\it CombCascade}. The results are shown in Fig.~\ref{fig:movielens} and demonstrate the advantage to involve context information in real applications.

@@

In the last experiment, we evaluated $C^3$-UCB on a problem of network routing with the {\it RocketFuel} dataset \cite{spring2004measuring}.

The ground set $E$ is the set of links in the network. Before learning, the environment randomly chooses a $d$-dimensional vector $\theta_* \in [0,1]^d$. At each time $t$, a pair of source and destination nodes are randomly chosen and the feasible action set $\cS_t$ at time $t$ is all simple paths between the source and destination nodes. Any edge $a$ in the set $\cS$ is assigned with a random $d$-dimensional context vector $x_{t,a}$. Notice here both $\theta$ and $x$ have been processed like section \ref{sec:ExpSyn} so that $\theta_*^{\top}x \in [0,1]$. The weight for each edge $a$ is a sample from Bernoulli distribution with mean $\theta_*^{\top}x_{t,a}$. Then the learning agent recommends a feasible path $A = (a_1,...,a_{|A|})$ between source and destination nodes to maximize the expected reward in the conjunctive objective. We experimented with different position discounts and different networks. The results are shown in Fig.~\ref{fig:network} (a) (b).

We also experimented on the network with latency, where the latency on an edge is an exponential random variable with mean $\theta_{\ast}^{\top}x$ and the reward is $1$ if the latency is less than a tolerance $\tau = 0.75$. The comparison of rewards of our algorithm with {\it CombCascade} is shown in Fig.~\ref{fig:network} (c).

We also experimented on the influence of different $\gamma$'s to the regret. In recommendations, the position discounts represent users' satisfaction and the learning agent may not know its true value. Suppose the true position discount $\gamma = 0.9$ and the learning agent sets $\gamma$ by her experience. We experiment on different setting $\gamma \in \{0.8, 0.82, \ldots, 0.98, 1.0\}$ with the true criterion $\gamma = 0.9$. The regrets with different $\gamma$'s are illustrated in Figure~\ref{fig:network}(d). The setting $\gamma$ is more close to the true $\gamma$, the regret is tending to be smaller. This experiment shows when there are position discounts in real application, it is better to represent this information to the model.
